04_analysis

### 2023-05-15 04_analysis.Rmd
updated with ALL buoys

---
title: "04_analysis"
author: "Marina Bozinovic"
date: "2023-05-11"
output: html_document
---

Required Packages
```{r}
# add packages here!!
```

@@

Get avg daily environmental variables
```{r}
means_08 <- tracks_SS_08 %>%
  group_by(as.Date(UTC))%>%
  summarize(sst_mean = mean(sst_mean),
            dist_slope = mean(dist2slope),
            depth = mean(depth),
            slope = mean(slope),
            dist_shore = mean(dist2shore),
            chlorophyll_mean = mean(chlorophyll_mean),
            ssh_mean = mean(ssh_mean),
            mldDepth = mean(mldDepth),
            mldTemp = mean(mldTemp),
            ttDepth = mean(ttDepth),
            ttTemp = mean(ttTemp),
            temp400 = mean(temp400),
            sal400 = mean(sal400),
            BWpresence = sum(BWpresence))
```

### Exploratory Analysis ######

@@

```{r}
# Drift 8 - all TOL bands, facet wrap
pivot <- tracks_SS_08 %>%
  pivot_longer(cols = starts_with("TOL") | starts_with("BB"), names_to = "SL", values_to = "Hz")    # Pivot longer for histogram facet grid

ggplot(data = pivot, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 3, color = "black") +
  theme_classic() +
  ggtitle("Distribution of third octave level intensities: Drift 8") +
  labs(x = "Sound Intensity (dB)") +
  guides(fill = guide_legend(title = "Beaked Whale Presence")) +
  scale_fill_manual(values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.7,0.8)) +
  facet_grid(~SL)

# Drift 10 - all TOL bands, facet grid
pivot2 <- tracks_SS_10 %>%
  pivot_longer(cols = starts_with("TOL")| starts_with("BB"), names_to = "SL", values_to = "Hz")    # Pivot longer for histogram facet grid

ggplot(data = pivot2, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 3, color = "black") +
  theme_classic() +
  ggtitle("Distribution of third octave level intensities: Drift 10") +
  scale_x_continuous(name = "Sound Intensity (dB)", breaks=seq(50,110,20)) +
  guides(fill = guide_legend(title = "Beaked Whale Presence")) +
  scale_fill_manual(values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.6,0.76)) +
  facet_grid(~SL)

# Drift 8 - TOL bands 125 and 2000 only
pivot3 <- tracks_SS_08 %>%
  dplyr::select(-TOL_63, -TOL_20000, -TOL_5000) %>%
  pivot_longer(cols = starts_with("TOL"), names_to = "TOL", values_to = "Hz")    # Pivot longer for histogram facet grid

TOL.labs <- c(`TOL_125` = "TOL at 125 Hz", 
              `TOL_2000` = "TOL at 2000 Hz")
```

```{r drift-8-histogram}
ggplot(data = pivot3, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 1, color = "black") +
  theme_classic() +
  ggtitle("Distribution of Third-Octave Level (TOL) Intensities: Buoy 8") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_manual(name = "Beaked Whale Presence", 
                    labels = c("0" = "Absent", "1" = "Present"),
                    values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.6,0.7)) +
  facet_wrap(~TOL, labeller = as_labeller(TOL.labs))
```

```{r}
# Drift 10 - TOL bands 125 and 2000 only
pivot4 <- tracks_SS_10 %>%
  dplyr::select(-TOL_63, -TOL_20000, -TOL_5000) %>%
  pivot_longer(cols = starts_with("TOL"), names_to = "TOL", values_to = "Hz")    # Pivot longer for histogram facet grid

ggplot(data = pivot4, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 1, color = "black") +
  theme_classic() +
  ggtitle("Distribution of Third-Octave Level (TOL) Intensities: Buoy 10") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_manual(name = "Beaked Whale Presence", 
                    labels = c("0" = "Absent", "1" = "Present"),
                    values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.6,0.7)) +
  facet_wrap(~TOL, labeller = as_labeller(TOL.labs))

TOL.labs <- c(`TOL_125` = "TOL at 125 Hz", 
              `TOL_2000` = "TOL at 2000 Hz")
```

@@

```{r}
# SST vs TOL 2000
ggplot(data = tracks_SS_08, aes(x = sst_mean, y = `TOL_2000`)) +
  geom_point(aes(color = species)) + scale_color_discrete(na.translate=FALSE)

# SST over time
ggplot(tracks_SS_08, aes(x = UTC, y = sst_mean)) +
  geom_line() +
  geom_point(aes(color = species)) +
  scale_color_discrete(na.translate=FALSE)
  #geom_smooth(method = "loess")

# Depth vs TOL 125
ggplot(data = tracks_SS_08, aes(x = depth, y = `TOL_125`)) +
  geom_line() +
  geom_point(aes(color = species)) +
  scale_color_discrete(na.translate=FALSE)

#Mean depth over time
ggplot(means_08, aes(x = `as.Date(UTC)`, y = depth*(-1))) +
  geom_line()

# Chlorophyll vs TOL 125
ggplot(data = tracks_SS_08, aes(x = chlorophyll_mean, y = `TOL_125`)) +
  geom_point(aes(color = species))

# CHl over time
ggplot(tracks_SS_08, aes(x = UTC, y = chlorophyll_mean)) +
  geom_line() +
  geom_point(aes(color = species)) +
  scale_color_discrete(na.translate=FALSE)
  #geom_smooth(method = "loess")

# Mixed Layer Depth vs TOL125 with species
ggplot(data = tracks_SS_08, aes(x = mldDepth, y = `TOL_125`)) +
  geom_point(aes(color = species))

# Mixed Layer Depth over time
ggplot(tracks_SS_08, aes(x = UTC, y = mldDepth*(-1))) +
  geom_line() +
  geom_point(aes(color = species)) +
  scale_color_discrete(na.translate=FALSE)

ggplot(data = tracks_SS_08, aes(x = slope, y = `TOL_5000`)) +
  geom_point(aes(color = species))
```

Facet Grids for Env. Variables
```{r}
#For surface/depth variables USING DAILY MEAN VARIABLES
# See above for means_08 derivation
track08Long <- means_08 %>%
  pivot_longer(cols = sst_mean:ssh_mean,
               names_to="parameter",
               values_to="value") %>%
  filter(parameter %in% c("sst_mean", "dist_shore", "chlorophyll_mean", "ssh_mean")) %>%
  rename(Date = `as.Date(UTC)`)

p <- ggplot(data = track08Long, aes(x = Date, y = value)) +
  geom_line()
 # geom_point(aes(size = BWpresence))   ### How best to represent number of BW detections per day?

@@

# For MLD/Thermocline variables 
track08Long2 <- tracks_SS_08 %>%
  pivot_longer(cols = mldDepth:sal400,
               names_to="parameter",
               values_to="value") %>%
  filter(parameter %in% c("mldDepth", "mldTemp", "ttDepth", "ttTemp", "temp400", "sal400")) %>%
    rename(Date = UTC) %>%
    subset(species!="?BW")

q <- ggplot(data = track08Long2, aes(x = Date, y = value)) +
  geom_line() + geom_point(aes(color = species)) + scale_color_discrete(na.translate=FALSE)
q + facet_grid(parameter ~ ., scales = "free_y")

@@

track08Long3 <- tracks_SS_08 %>%
  pivot_longer(cols = temp400:sal400,
               names_to="parameter",
               values_to="value") %>%
  filter(parameter %in% c("temp400", "sal400")) %>%
  rename(`Date (2018)` = UTC)

q <- ggplot(data = track08Long3, aes(x = `Date (2018)`, y = value)) +
  geom_line() + 
  geom_point(aes(color = species)) + 
  scale_color_discrete(na.translate=FALSE, labels = c("ZC" = "Cuvier's", "BW43" = "BW43", "BW" = "BW"))
q + facet_grid(parameter ~ ., scales = "free_y", labeller = as_labeller(env.labs))
```

Co-occurrence plots of drift 8 and vessels 
```{r}
#### 09/04/18 co-occurrence of vessel and drift within 10 km
track08_09.04 <- tracks_SS_08 %>%
  filter(UTC >= "2018-09-04 00:00:00" & UTC <= "2018-09-05 00:00:00")

@@

BW Occurrences
```{r}
# Filter Drift 8 by BW presence and absence
BW1 <- tracks_SS_08 %>%
  filter(BWpresence == 1)
BW0 <- tracks_SS_08 %>%
  filter(BWpresence == 0)

@@

ggplot() + 
  geom_histogram(data = BW0, aes(TOL_5000), binwidth = 0.5, fill = "black", color = "white") +
  geom_histogram(data = BW1, aes(TOL_5000), binwidth = 0.5, fill = "yellow", color = "black")

# Percent of time BWs are detected
print(paste(nrow(BW1) / nrow(tracks_SS_08) * 100,"%"))

# Most abundant BW species when whale is detected (remove NA values)
tracks_SS_08 %>%
  filter(!is.na(species)) %>%
  ggplot() + geom_bar(aes(species))

# Sound levels BW are exposed to by species
ZC <- tracks_SS_08 %>%
  filter(species == "ZC")

BW43 <- tracks_SS_08 %>%
  filter(species == "BW43")

print(max(ZC$TOL_125))
print(max(BW43$TOL_125))

@@

Density Plots - Species at TOL sound levels
```{r}
# Broadband
tracks_SS_08 %>%
  filter(!is.na(species)) %>%
  ggplot(aes(`BB_20.24000`, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("Broadband") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 63
tracks_SS_08 %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_63, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("63 Hz") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 125
tracks_SS_08 %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_125, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("125 Hz") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 2000
tracks_SS_08 %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_2000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("2000 Hz") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 5000
tracks_SS_08 %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_5000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("5000 Hz") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 20,000
tracks_SS_08 %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_20000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("20,000 Hz") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))
```

Histograms with BW presence/absence and environmental variables
```{r}
# SST
ggplot(data = tracks_SS_08, aes(sst_mean, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.1, color = "black")

# Depth
ggplot(data = tracks_SS_08, aes(depth, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 10, color = "black")

# Bottom slope
ggplot(data = tracks_SS_08, aes(slope, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.005, color = "black")

# SSH
ggplot(data = tracks_SS_08, aes(ssh_mean, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.005, color = "black")

# Chlorophyll
ggplot(data = tracks_SS_08, aes(chlorophyll_mean, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.005, color = "black")

# Mixed layer depth Depth
ggplot(data = tracks_SS_08, aes(mldDepth, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.25, color = "black")

# Mixed layer depth Temperature
ggplot(data = tracks_SS_08, aes(mldTemp, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.1, color = "black")

# Thermocline Depth
ggplot(data = tracks_SS_08, aes(ttDepth, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.25, color = "black")

#Thermocline temperature
ggplot(data = tracks_SS_08, aes(ttTemp, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.05, color = "black")

# Temperature at 400m
ggplot(data = tracks_SS_08, aes(temp400, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.01, color = "black")

# Salinity at 400 m
ggplot(data = tracks_SS_08, aes(sal400, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.001, color = "black")
```

@@

Create sf objects and basemaps for maps (not needed for environmental variables)
```{r}
whales_8_10sf <- BWsf %>%                            
dplyr::filter(station == "8" | station == "10")

@@

Map of whale detections
```{r}
# Whale detections
ggplot() +
  geom_point(data = whales8, mapping = aes(x=long, y=lat), shape = 21, color = "black", fill = "green4", size = 4) +
  geom_point(data = whales10, mapping = aes(x=long, y=lat), shape = 21, color = "black", fill = "green4", size = 4) +
  scale_x_continuous(limits = c(-130,-122)) + 
  scale_y_continuous(limits = c(35,39)) +
  annotation_scale(location = "br", width_hint = 0.2, plot_unit = "km") +
  theme(panel.border = element_rect(fill = NA),
        axis.text = element_blank(),
        axis.ticks = element_blank(), 
        axis.title = element_blank(),
        panel.background = element_blank()
        )
#ggsave("BWdetect.png")

@@

# Soundscape + bathymetry
whales08_sf <- BWsf %>%                            
  dplyr::filter(station == "8")

>>>>

---
title: "04_analysis"
author: "Marina Bozinovic"
date: "2023-05-11"
output: html_document
---
# Exploratory Analysis and Modeling for all 13 buoys

Required Packages
```{r}
library(tidyverse)
```

@@

Bind ALL DRIFTS into one dataframe
```{r}
allDrifts <- rbind(fulltrack07, fulltrack08, fulltrack10, fulltrack12, fulltrack13, fulltrack14, 
                   fulltrack16, fulltrack18, fulltrack19, fulltrack20, fulltrack21, fulltrack22, fulltrack23)
```

@@

```{r all-drifts-all-freq-histogram}
# create list of fulltrack dfs from global environment to loop through
dlist <- grep(pattern = "fulltrack", names(.GlobalEnv), value = T)

# all Drifts - all TOL bands, facet wrap
for (ft in dlist) {
pivot1 <- get(ft) %>%
  pivot_longer(cols = starts_with("TOL") | starts_with("BB"), names_to = "SL", values_to = "Hz")    # Pivot longer for histogram facet grid

print(ggplot(data = pivot1, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 3, color = "black") +
  theme_classic() +
  ggtitle(paste0("Distribution of third octave level intensities: Drift ", substr(ft,10,11))) +
  labs(x = "Sound Intensity (dB)") +
  guides(fill = guide_legend(title = "Whale Presence")) +
  scale_fill_manual(values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.7,0.8)) +
  facet_grid(~SL))
}

# Single plot for ALL DRIFTS
pivot3 <- allDrifts %>%
  pivot_longer(cols = starts_with("TOL") | starts_with("BB"), names_to = "SL", values_to = "Hz")    # Pivot longer for histogram facet grid

ggplot(data = pivot3, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 3, color = "black") +
  theme_classic() +
  ggtitle("Distribution of third octave level intensities: All Drifts Combined") +
  labs(x = "Sound Intensity (dB)") +
  guides(fill = guide_legend(title = "Whale Presence")) +
  scale_fill_manual(values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.7,0.8)) +
  facet_grid(~SL)

## ignore??
```{r drift-8-histogram}
ggplot(data = pivot3, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 1, color = "black") +
  theme_classic() +
  ggtitle("Distribution of Third-Octave Level (TOL) Intensities: Buoy 8") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_manual(name = "Beaked Whale Presence", 
                    labels = c("0" = "Absent", "1" = "Present"),
                    values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.6,0.7)) +
  facet_wrap(~TOL, labeller = as_labeller(TOL.labs))
```

```{r all-drifts-two-freq-histogram}
# All Drifts - TOL bands 125 and 2000 only

TOL.labs <- c(`TOL_125` = "TOL at 125 Hz", 
              `TOL_2000` = "TOL at 2000 Hz") #creates label names for plots below

for (ft in dlist) {
pivot2 <- get(ft) %>%
  dplyr::select(-TOL_63, -TOL_20000, -TOL_5000) %>%
  pivot_longer(cols = starts_with("TOL"), names_to = "TOL", values_to = "Hz")    # Pivot longer for histogram facet grid

print(ggplot(data = pivot2, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 1, color = "black") +
  theme_classic() +
  ggtitle(paste0("Distribution of Third-Octave Level (TOL) Intensities: Drift ", substr(ft,10,11))) +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_manual(name = "Whale Presence", 
                    labels = c("0" = "Absent", "1" = "Present"),
                    values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.6,0.7)) +
  facet_wrap(~TOL, labeller = as_labeller(TOL.labs)))
}

# Single plot for ALL DRIFTS combined
pivot4 <- allDrifts %>%
  dplyr::select(-TOL_63, -TOL_20000, -TOL_5000) %>%
  pivot_longer(cols = starts_with("TOL"), names_to = "TOL", values_to = "Hz")    # Pivot longer for histogram facet grid

ggplot(data = pivot4, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 1, color = "black") +
  theme_classic() +
  ggtitle("Distribution of Third-Octave Level (TOL) Intensities: All Drifts Combined") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_manual(name = "Whale Presence", 
                    labels = c("0" = "Absent", "1" = "Present"),
                    values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.6,0.7)) +
  facet_wrap(~TOL, labeller = as_labeller(TOL.labs))
```

Timeline of drifts in ocean
```{r}
ggplot(data = allDrifts, aes(x = UTC, y = station*(-1))) + geom_point(aes(color = factor(station)))

#Note the break in drift 20 - I think it wasn't transmitting for several days (11/15-11/19)
```

@@

```{r}
# SST vs TOL 2000
ggplot(data = allDrifts, aes(x = sst_mean, y = `TOL_2000`)) +
  geom_point(aes(color = species)) + scale_color_discrete(na.translate=FALSE)

# SST over time (strange plot)
ggplot(allDrifts, aes(x = UTC, y = sst_mean)) +
  geom_line() +
  geom_point(aes(color = species)) +
  scale_color_discrete(na.translate=FALSE)
  #geom_smooth(method = "loess")

# Depth vs TOL 125
ggplot(data = allDrifts, aes(x = depth, y = `TOL_125`)) +
  geom_line() +
  geom_point(aes(color = species)) +
  scale_color_discrete(na.translate=FALSE)

#Mean depth over time
#ggplot(means_08, aes(x = `as.Date(UTC)`, y = depth*(-1))) +
#  geom_line()

# Chlorophyll vs TOL 125
ggplot(data = allDrifts, aes(x = chlorophyll_mean, y = `TOL_125`)) +
  geom_point(aes(color = species)) + scale_color_discrete(na.translate=FALSE)

# CHl over time
ggplot(allDrifts, aes(x = UTC, y = chlorophyll_mean)) +
  geom_line() +
  geom_point(aes(color = species)) +
  scale_color_discrete(na.translate=FALSE)
  #geom_smooth(method = "loess")

# Mixed Layer Depth vs TOL125 with species
ggplot(data = allDrifts, aes(x = mldDepth, y = `TOL_125`)) +
  geom_point(aes(color = species)) + scale_color_discrete(na.translate=FALSE)

# Mixed Layer Depth over time
ggplot(allDrifts, aes(x = UTC, y = mldDepth*(-1))) +
  geom_line() +
  geom_point(aes(color = species)) +
  scale_color_discrete(na.translate=FALSE)

ggplot(data = allDrifts, aes(x = `TOL_125`, y = bathy_slope)) +
  geom_point(aes(color = species)) + scale_color_discrete(na.translate=FALSE)

Facet Grids for Env. Variables
```{r}
#Get avg daily environmental variables
means <- allDrifts %>%
  group_by(as.Date(UTC))%>%
  summarize(sst_mean = mean(sst_mean),
            dist_slope = mean(dist2slope),
            depth = mean(depth),
            slope = mean(bathy_slope),
            dist_shore = mean(dist2shore),
            chlorophyll_mean = mean(chlorophyll_mean),
            ssh_mean = mean(ssh_mean),
            mldDepth = mean(mldDepth),
            mldTemp = mean(mldTemp),
            ttDepth = mean(ttDepth),
            ttTemp = mean(ttTemp),
            temp400 = mean(temp400),
            sal400 = mean(sal400),
            curlmean = mean(curl_mean),
            Wh_presence = sum(BWpresence))
```

```{r}
#For surface/depth variables USING DAILY MEAN VARIABLES
trackLong <- means %>%
  pivot_longer(cols = sst_mean:ssh_mean,
               names_to="parameter",
               values_to="value") %>%
  filter(parameter %in% c("sst_mean", "dist_shore", "chlorophyll_mean", "ssh_mean")) %>%
  rename(Date = `as.Date(UTC)`)

p <- ggplot(data = trackLong, aes(x = Date, y = value)) +
  geom_line()

@@

# For MLD/Thermocline variables 
trackLong2 <- allDrifts %>%
  pivot_longer(cols = mldDepth:sal400,
               names_to="parameter",
               values_to="value") %>%
  filter(parameter %in% c("mldDepth", "mldTemp", "ttDepth", "ttTemp", "temp400", "sal400")) %>%
    rename(Date = UTC) %>%
    subset(species!="?BW")

q <- ggplot(data = trackLong2, aes(x = Date, y = value)) +
  geom_line() + geom_point(aes(color = species)) + scale_color_discrete(na.translate=FALSE)
q + facet_grid(parameter ~ ., scales = "free_y")

@@

trackLong3 <- allDrifts %>%
  pivot_longer(cols = temp400:sal400,
               names_to="parameter",
               values_to="value") %>%
  filter(parameter %in% c("temp400", "sal400")) %>%
  rename(`Date (2018)` = UTC)

q <- ggplot(data = trackLong3, aes(x = `Date (2018)`, y = value)) +
  #geom_line() + 
  geom_point(aes(color = species)) + 
  scale_color_discrete(na.translate=FALSE, labels = c("ZC" = "Cuvier's", "BW43" = "BW43", "BW" = "BW"))
q + facet_grid(parameter ~ ., scales = "free_y", labeller = as_labeller(env.labs))
```

NEEDS FIXING/LOOP/AIS DATA FOR OTHER DRIFTS
Co-occurrence plots of drift 8 and vessels 
```{r}
#### 09/04/18 co-occurrence of vessel and drift within 10 km
track08_09.04 <- tracks_SS_08 %>%
  filter(UTC >= "2018-09-04 00:00:00" & UTC <= "2018-09-05 00:00:00")

@@

IS THIS HELPFUL??
Whale Occurrences
```{r}
# Filter Drift 8 by Whale presence and absence
W1 <- tracks_SS_08 %>%
  filter(BWpresence == 1)
W0 <- tracks_SS_08 %>%
  filter(BWpresence == 0)

@@

ggplot() + 
  geom_histogram(data = W0, aes(TOL_5000), binwidth = 0.5, fill = "black", color = "white") +
  geom_histogram(data = W1, aes(TOL_5000), binwidth = 0.5, fill = "yellow", color = "black")

# Percent of time BWs are detected
print(paste(nrow(W1) / nrow(tracks_SS_08) * 100,"%"))

# Most abundant whale species when whale is detected (remove NA values)
allDrifts %>%
  filter(!is.na(species)) %>%
  ggplot() + geom_bar(aes(species))

# Sound levels whales are exposed to by species
ZC <- allDrifts %>%
  filter(species == "ZC")

BW43 <- allDrifts %>%
  filter(species == "BW43")

BB <- allDrifts %>%
  filter(species == "BB")

BWC <- allDrifts %>%
  filter(species == "BWC")

BW37V <- allDrifts %>%
  filter(species == "BW37V")

PM <- allDrifts %>%
  filter(species == "PM")

print(max(ZC$TOL_125))
print(max(BW43$TOL_125))
print(max(BB$TOL_125))
print(max(BWC$TOL_125))
print(max(BW37V$TOL_125))
print(max(PM$TOL_125))

@@

Density Plots - Species at TOL sound levels
```{r density-plots}
# Broadband
for (ft in dlist) {
plt <- get(ft) %>%
  filter(!is.na(species))

  print(ggplot(data = plt, aes(`BB_20.24000`, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle(paste0("Broadband Drift ", substr(ft, 10, 11))) +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43")))
}

# Broadband SINGLE PLOT ALL DRIFTS
allDrifts %>%
  filter(!is.na(species)) %>%
  ggplot(aes(`BB_20.24000`, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("Broadband ALL Drifts Combined") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 63
for (ft in dlist) {
plt <- get(ft) %>%
  filter(!is.na(species))

  print(ggplot(data = plt, aes(TOL_63, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle(paste0("TOL 63, Drift ", substr(ft, 10, 11))) +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43")))
}

# TOL 63 SINGLE PLOT ALL DRIFTS
allDrifts %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_63, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("TOL 63 ALL Drifts Combined") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 125
for (ft in dlist) {
plt <- get(ft) %>%
  filter(!is.na(species))

  print(ggplot(data = plt, aes(TOL_125, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle(paste0("TOL 125, Drift ", substr(ft, 10, 11))) +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43")))
}

# TOL 125 SINGLE PLOT ALL DRIFTS
allDrifts %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_125, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("TOL 125 ALL Drifts Combined") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 2000
for (ft in dlist) {
plt <- get(ft) %>%
  filter(!is.na(species))

  print(ggplot(data = plt, aes(TOL_2000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle(paste0("TOL 2000, Drift ", substr(ft, 10, 11))) +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43")))
}

# TOL 2000 SINGLE PLOT ALL DRIFTS
allDrifts %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_2000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("TOL 2000 ALL Drifts Combined") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 5000
for (ft in dlist) {
plt <- get(ft) %>%
  filter(!is.na(species))

  print(ggplot(data = plt, aes(TOL_5000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle(paste0("TOL 5000, Drift ", substr(ft, 10, 11))) +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43")))
}

# TOL 5000 SINGLE PLOT ALL DRIFTS
allDrifts %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_5000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("TOL 5000 ALL Drifts Combined") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 20,000
for (ft in dlist) {
plt <- get(ft) %>%
  filter(!is.na(species))

  print(ggplot(data = plt, aes(TOL_20000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle(paste0("TOL 20000, Drift ", substr(ft, 10, 11))) +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43")))
}

# TOL 20,000 SINGLE PLOT ALL DRIFTS
allDrifts %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_20000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("TOL 20,000 ALL Drifts Combined") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))
```

ARE THESE HELPFUL??
Histograms with whale presence/absence and environmental variables. Doesnt differentiate by species though
```{r}
# SST
ggplot(data = allDrifts, aes(sst_mean, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.1, color = "black")

# Depth
ggplot(data = allDrifts, aes(depth, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 100, color = "black")

# Bottom slope
ggplot(data = allDrifts, aes(bathy_slope, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 1, color = "black")

# SSH
ggplot(data = allDrifts, aes(ssh_mean, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.005, color = "black")

# Chlorophyll
ggplot(data = allDrifts, aes(chlorophyll_mean, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.05, color = "black")

#Wind Curl
ggplot(data = allDrifts, aes(curl_mean, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.0000005, color = "black")

# Mixed layer depth Depth
ggplot(data = allDrifts, aes(mldDepth, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.25, color = "black")

# Mixed layer depth Temperature
ggplot(data = allDrifts, aes(mldTemp, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.1, color = "black")

# Thermocline Depth
ggplot(data = allDrifts, aes(ttDepth, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.75, color = "black")

#Thermocline temperature
ggplot(data = allDrifts, aes(ttTemp, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.1, color = "black")

# Temperature at 400m
ggplot(data = allDrifts, aes(temp400, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.03, color = "black")

# Salinity at 400 m
ggplot(data = allDrifts, aes(sal400, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.005, color = "black")
```

@@

Create sf objects and basemaps for maps (not needed for environmental variables)
```{r}
whales_8_10sf <- Wsf %>%                            
dplyr::filter(station == "8" | station == "10")

@@

Map of whale detections
```{r}
# Whale detections
ggplot() +
  geom_point(data = whales8, mapping = aes(x=long, y=lat), shape = 21, color = "black", fill = "green4", size = 4) +
  geom_point(data = whales10, mapping = aes(x=long, y=lat), shape = 21, color = "black", fill = "green4", size = 4) +
  scale_x_continuous(limits = c(-130,-122)) + 
  scale_y_continuous(limits = c(35,39)) +
  annotation_scale(location = "br", width_hint = 0.2, plot_unit = "km") +
  theme(panel.border = element_rect(fill = NA),
        axis.text = element_blank(),
        axis.ticks = element_blank(), 
        axis.title = element_blank(),
        panel.background = element_blank()
        )
#ggsave("Wdetect.png")

@@

# Soundscape + bathymetry
whales08_sf <- Wsf %>%                            
  dplyr::filter(station == "8")
